\mytitle{}

### 2018-11-06 paper/paper.tex
Section V drafted

\subsection{Motivation}
%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
%?? Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows. This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e. an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
Moving beyond the overt benefits of choosing RL-based defences for coping with non-stationary problems, we believe that there are concrete reasons for their use here.
We have seen that for other domains in particular, misclassification is a serious problem, which can introduce \emph{collateral damage} in the context of DDoS prevention.
In theory, the feedback-loop-like model allows us to monitor flows \emph{after} an action is taken to allow forgiveness of mistakenly punished flows.
This does rely upon the ability to take a flow-by-flow view of the state space, but if we can combine knowledge of current state with the last applied action, then perhaps a flow which falls off identically to a legitimate flow can be rescued.

@@

We employ the reward functions introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, we make use of their global reward as a performance metric and their \emph{coordinated team learning} (CTL) reward for training.
These assume the presence of a classifier, $g(\cdot)$, which estimates the volume of `legitimate traffic' received at any point in the network.
At time $t$ an agent $v$ (with team leader $l_v$) receives either $R_{t}^{\mathit{Global}}$ or $R_{t}^{\mathit{CTL}}$:
\begin{subequations}
\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
\begin{gather}
\cond{f} = [\load{s} > U_s],\\
\cond{c}{(v)} = [\load{l_v} > U_s/k],\\
R_{t}^{\mathit{Global}} = (1 - \cond{f}) \frac{g(\load{s})}{g(\load{\mathit{total}})} - \cond{f},\label{eqn:reward-rt}\\
R_{t}^{\mathit{CTL}}(v) = (1 - \cond{f}\cond{c}{(v)}) \frac{g(\load{l_v})}{g(\load{\mathit{team}})} \allowbreak - \cond{f} \cond{c}{(v)},\label{eqn:reward-rctl}
\end{gather}
where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $v$'s team, respectively.\label{eqn:reward}
\end{subequations}

@@

\subsection{TCP Multiplexing and Fallback Behaviour}\label{sec:tcp-multiplexing-and-fallback-behaviour}

@@

These results are of particular importance due to TCP's prevalence within the Internet.
The protocol is known to carry \SIrange{79}{94}{\percent} of packets, corresponding to \SIrange{89}{98}{\percent} of data transferred by volume \cite{DBLP:conf/saint/ZhangDJC09}.
As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development.

@@

{
\newcommand{\load}[2]{\operatorname{load}^{#2}_{t}(#1)}
\newcommand{\uload}[1]{\load{#1}{\uparrow}}
\newcommand{\dload}[1]{\load{#1}{\downarrow}}
\newcommand{\bload}[1]{\load{#1}{\updownarrow}}
\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
%\fakepara{Reward function directionality}
The reward functions, as defined, do not take traffic direction into account.
We redefine these to identify overload states using both upstream and downstream loads, while allowing customisation of which direction is chosen for protection.
Denoting the upstream, downstream and combined loads $\uload{s}, \dload{s}, \bload{s}$, we modify \cref{eqn:reward}:
\begin{subequations}
\begin{gather}
\cond{f} = [\max(\uload{s}, \dload{s}) > U_s],\\
\cond{c}{(v)} = [\max(\uload{l_v}, \dload{l_v}) > U_s/k],
\end{gather}
\end{subequations}
replacing $\load{t}{}$ in \crefrange{eqn:reward-rt}{eqn:reward-rctl} with whichever directional load is prioritised.
We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, though we expect that $\bload{\cdot}$ would be the most suitable for general deployment or heterogeneous traffic patterns.
}

@@

An important question to ask while exploring the design space of RL agents is whether (and how) domain knowledge can improve the performance or learning rate of a defence system.
This often comes as a trade-off: expressive power and unfettered control, weighed against the smaller action spaces afforded by existing insight or structure.
We present a second model, relying on the measurements of \textcite{DBLP:conf/ndss/KangGS16} which suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
While it is not stated whether this behaviour applies to amplification attacks, we expect this to remain the case; the upload rate of any amplifier will be fixed at its upper bound given that the upstream links of any connected bots are saturated.
Moreover, bots receive no feedback on the effective throughput of their attack flows.

How might we apply their observations within the RL paradigm, while tailoring them towards non-ISP deployment?
Our approach is to constrain how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, reducing the action set to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ in single steps.
To enable temporary bandwidth expansion in resource constrained deployments, every flow is initially placed under light packet drop ($p=0.05$, or state 1), corresponding to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and non-congestion-aware traffic as required.
We combine this with the enhancements described in \cref{sec:marl-plus}.
%?? Talk about SPF -- Influenced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}. Why? To demonstrate the design of a model which includes some aspects of real-world knowledge, and to see whether making an agent less ``free-form'' might allow us to trade power for learning speed (smaller action space etc.). Also, because we might be able to use its insights in an environ where we don't have spare bandwidth? Such a model means we need an agents need to ``plan ahead'', i.e. have $\gamma \ne 0$. We found $\gamma = 0.8$ to be the most effective choice. We also use most of the techniques from above.

This model is significantly different when considering how to tune the RL algorithm's parameters: each agent now requires the capability to plan ahead, since they may no longer choose $p$ directly from any state.
This entails choosing some discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
We found $\gamma = 0.8$ to be the most effective choice during exploratory testing.

@@

\begin{figure*}
	\centering
	\begin{subfigure}{0.32\linewidth}
		\includegraphics[width=\linewidth]{../plots/ftprep-good}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\includegraphics[width=\linewidth]{../plots/ftprep-good-2}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\includegraphics[width=\linewidth]{../plots/ftprep-good-3}
	\end{subfigure}
	\caption{
		Features: how good are they really? These are for UDP. As you can see, Most things are pretty average (except for \emph{Mean IAT}, which dominates).
	\label{fig:udp-feature-plots}
}
\end{figure*}

@@

	\caption{
		Features: how good are they really (part 2)? These are for TCP. These are less conclusive...
		\label{fig:tcp-feature-plots}
	}
\end{figure*}

@@

	\caption{
		Features: how good are they really (part 3)? These are for TCP, but capped. These are less conclusive...
		\label{fig:tcp-cap-feature-plots}
	}
\end{figure*}

?? Mention Colin/Qianru's concerns about Quic, IPv6. At this point, src/dst IP are the only ``traditional'' characteristics that can be relied upon. Even then, IPv6 can have a different IP for each application ON A HOST. Port number meaningless...

?? NEED TO SAY SOMEWHERE: Baselines use TC for establishing interleaving/buffers filling etc., when MARL+ or SPF are in play then disabling TC (thus allowing illegal states, with \num{-1} reward marking them as such---``negative reinforcement'') may accelerate learning. We have a choice---but the results show that for TCP we WANT TC enabled... For UDP, we actually get a slight degradation in performance. This suggests a kind of ``banding'', perhaps?

?? What do we want?

?? note---probably put in the experiments that Simon suggested at this point as justification for why we chose what we did...

\textbf{Global state}---what came before (load statistics from various network locations)

\textbf{Local state}---per-source data. Specifically, right now I'm using:

?? \emph{IP}

?? \emph{Last action taken}: encodes the notion of current belief of flow maliciousness (default at 0). Potentially allows forgiveness: a mistakenly malicious flow's falloff after punishment could be used for such purpose (if we're lucky...)

?? \emph{Flow duration, size}

?? \emph{Correspondence ratio}: ratio (in bytes) between upstream/downstream traffic for a src/dst pair. (Cite the work that suggested this) $C_X = \min(\mathit{in}, \mathit{out})/\max(\mathit{in}, \mathit{out})$

?? \emph{Change in send/recv rates} since last action/check.

?? \emph{Mean inter-arrival time}.

?? \Cref{fig:udp-feature-plots} shows the effectiveness of each for UDP (resp.\ \cref{fig:udp-feature-plots} for TCP).

\begin{table}
	\centering
	\caption{Tile coding windows for each feature.}

@@

\subsection{What's done with these?}

?? Tile coded together, action taken per-flow every timestep.

?? TODO: everything

?? Explicitly mention \emph{why} we don't make use of neural networks, and instead favour more classical function approximation. Latency concerns, energy usage, graphics/TPU/whatever hardware availability or lack thereof across switches/routers/etc.

?? NOTE: think about how I might need to change/alter the tile-coding when we add extra discriminative features... (i.e., are there better ways to handle covariance between properties of system/flow state, properly capture generalisation?)

?? What other consequences are there? We're training each agent on all the flows it's acting on which means slight theoretical changes (we have multiple experience traces, and each uses its last state + new state + new target value (reward) as an update to the \emph{current model}).

?? Have some more things worth testing now. FSA with transitions based on the theory underlying SPIFFY. 5 states (start in low pdrop) from \SIrange{0}{100}{\percent} drop, actions move up, down or hold on the level of drop. Why like this? Chosen deployment environments might not be able to reroute to get the requisite bandwidth epansion factor, so punish everyone to begin with.

@@

?? "We chose to take a separate tile-coding over each feature, ignoring combinatorial effects for now. Combining them all as one tile-coding led to poor performance due to an inability to generalise value across similar states."

@@

\fakepara{DDoS Prevention}
%?? Other recent work in the prevention of DDoS? (i.e., non-RL).
\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
%?? What does it do? detects state, not which flows are contributing...

@@

An \emph{indirect-control} RL problem, where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
%Indirect-control applications: link to all the juicy stuff in e.g. HotNets, data-driven networking, the works. \cite{DBLP:conf/hotnets/ValadarskySST17, DBLP:conf/hotnets/MaoAMK16}.
A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.

%?? Completely reword, yikes.
\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation, using the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}.
Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
To overcome the high latency of action computation via neural network, two agents are trained, handling aspects of short and long flows respectively.
The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.

@@

>>>>

\subsection{Motivation}\label{sec:motivation}
%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
%?? Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows. This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e. an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
Moving beyond the overt benefits of choosing RL-based defences for coping with non-stationary problems, we believe that there are concrete reasons for their use here.
We have seen that for other domains in particular, misclassification is a serious problem, which can introduce \emph{collateral damage} in the context of DDoS prevention.
In theory, the feedback-loop-like model allows us to monitor flows \emph{after} an action is taken to allow forgiveness of mistakenly punished flows.
This does rely upon the ability to take a flow-by-flow view of the state space, but if we can combine knowledge of current state with the last applied action, then perhaps a flow which falls off identically to a legitimate flow can be rescued.

@@

We employ the reward functions introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, we make use of their global reward as a performance metric and their \emph{coordinated team learning} (CTL) reward for training.
These assume the presence of a classifier, $g(\cdot)$, which estimates the volume of `legitimate traffic' received at any point in the network.
At time $t$ an agent $v$ (with team leader $l_v$) receives either $R_{t}^{\mathit{Global}}$ or $R_{t}^{\mathit{CTL}}$:
\begin{subequations}
\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
\begin{gather}
\cond{f} = [\load{s} > U_s],\\
\cond{c}{(v)} = \cond{f} \cdot{} \, [\load{l_v} > U_s/k],\\
R_{t}^{\mathit{Global}} = (1 - \cond{f}) \frac{g(\load{s})}{g(\load{\mathit{total}})} - \cond{f},\label{eqn:reward-rt}\\
R_{t}^{\mathit{CTL}}(v) = (1 - \cond{c}{(v)}) \frac{g(\load{l_v})}{g(\load{\mathit{team}})} \allowbreak - \cond{c}{(v)},\label{eqn:reward-rctl}
\end{gather}
where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $v$'s team, respectively.\label{eqn:reward}
\end{subequations}

@@

\subsection{TCP Multiplexing and Falloff Behaviour}\label{sec:tcp-multiplexing-and-fallback-behaviour}

@@

These results are of particular importance due to TCP's prevalence within the Internet.
The protocol is known to carry \SIrange{79}{94}{\percent} of packets, corresponding to \SIrange{89}{98}{\percent} of data transferred by volume \cite{DBLP:conf/saint/ZhangDJC09}.
As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development to properly handle traffic with such characteristics.

@@

{
\newcommand{\load}[2]{\operatorname{load}^{#2}_{t}(#1)}
\newcommand{\uload}[1]{\load{#1}{\uparrow}}
\newcommand{\dload}[1]{\load{#1}{\downarrow}}
\newcommand{\bload}[1]{\load{#1}{\updownarrow}}
\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
%\fakepara{Reward function directionality}
The reward functions, as defined, do not take traffic direction into account.
We redefine these to identify overload states using both upstream and downstream loads, while allowing customisation of which direction is chosen for protection.
Denoting the upstream, downstream and combined loads $\uload{s}, \dload{s}, \bload{s}$, we modify \cref{eqn:reward}:
\begin{subequations}
\begin{gather}
\cond{f} = [\max(\uload{s}, \dload{s}) > U_s],\\
\cond{c}{(v)} = \cond{f}{} \cdot{} \, [\max(\uload{l_v}, \dload{l_v}) > U_s/k],
\end{gather}
\label{eqn:reward-but-better}
\end{subequations}
replacing $\load{t}{}$ in \crefrange{eqn:reward-rt}{eqn:reward-rctl} with whichever directional load is prioritised.
We choose $\uload{\cdot}$ for our UDP-based models and $\dload{\cdot}$ for HTTP, though we expect that $\bload{\cdot}$ would be the most suitable for general deployment or heterogeneous traffic patterns.
}

%Explicitly mention \emph{why} we don't make use of neural networks, and instead favour more classical function approximation. Latency concerns, energy usage, graphics/TPU/whatever hardware availability or lack thereof across switches/routers/etc.
To make decisions cheaply and at low latency, we maintain the usage of semi-gradient Sarsa with tile coding, rather than making use of neural networks or more complicated function approximators.
Although this likely brings lower theoretical performance, there are more reasons than latency alone to favour classical methods; these include lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.

@@

An important question to ask while exploring the design space of RL agents is whether (and how) domain knowledge can improve the performance or learning rate of a defence system.
This often comes as a trade-off: expressive power and unfettered control, weighed against the smaller action spaces afforded by existing insight or structure.
We propose a second model, relying on the measurements of \textcite{DBLP:conf/ndss/KangGS16} which suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
While it is not stated whether this behaviour applies to amplification attacks, we expect this to remain the case; the upload rate of any amplifier will be fixed at its upper bound given that the upstream links of any connected bots are saturated.
Moreover, bots receive no feedback on the effective throughput of their attack flows.

How might we apply their observations within the RL paradigm, while tailoring them towards non-ISP deployment?
Our approach is to constrain how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, reducing the action set to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ in single steps.
To enable temporary bandwidth expansion in resource constrained deployments, every flow is initially placed under light packet drop ($p=0.05$, or state 1), corresponding to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and non-congestion-aware traffic as required.
We combine this with the enhancements described in \cref{sec:marl-plus}.

This model is significantly different when considering how to tune the RL algorithm's parameters: each agent now requires the capability to plan ahead, since they may no longer choose $p$ directly from any state.
This entails choosing some discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
We found $\gamma = 0.8$ to be the most effective choice during exploratory testing with UDP traffic.

@@

\begin{figure*}
	\centering
	\begin{subfigure}{0.32\linewidth}
		\includegraphics[width=\linewidth]{../plots/ftprep-good}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\includegraphics[width=\linewidth]{../plots/ftprep-good-2}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\includegraphics[width=\linewidth]{../plots/ftprep-good-3}
	\end{subfigure}
	\caption{
%		Features: how good are they really? These are for UDP. As you can see, Most things are pretty average (except for \emph{Mean IAT}, which dominates).
		Learned performance of Marl++ agents when benign traffic is UDP-like, using only a single feature as a basis for decisions.
		Mean IAT, inbound packet sizes, and global state offer the best predictive performance, while most features offer marginal advantage over the unprotected baseline.
		Here, the link's bandwidth limit is removed, and so agents automatically learn to keep bandwidth below $U_s$ by negative reinforcement---this offers a slight per-feature performance increase compared to when this explicit cap is present (not pictured).
		\label{fig:udp-feature-plots}
	}
\end{figure*}

@@

	\caption{
%		Features: how good are they really (part 2)? These are for TCP. These are less conclusive...
		Learned performance of Marl++ agents when benign traffic is TCP-like, using only a single feature as a basis for decisions (uncapped as in \cref{fig:udp-feature-plots}).
		Global state and Mean IAT still offer the greatest improvement above baseline, but it is observed that different optimal features are expressed between different traffic models.
		Packet-level statistics are considerably less effective for this class of traffic, with $\Delta$ Out Rate and Correspondence Ratio being proportionally better than they are alongside UDP.
		\label{fig:tcp-feature-plots}
	}
\end{figure*}

@@

	\caption{
%		Features: how good are they really (part 3)? These are for TCP, but capped. These are less conclusive...
		Learned performance of Marl++ agents when benign traffic is TCP-like, using only a single feature as a basis for decisions (with bandwidth capped physically).
		Although negative rewards (and thus negative reinforcement) cannot occur, all features achieve considerably better performance, a reversal of our observation for UDP traffic.
%		Global state and Mean IAT still offer the greatest improvement above baseline, but packet-level statistics are considerably less effective for this class of traffic.
		\label{fig:tcp-cap-feature-plots}
	}
\end{figure*}

The main element required to move to a per-source model is a feature set which offers high predictive power, such that behavioural differences are readily apparent to an agent.
Elaborating further on the statistics discussed in \cref{sec:motivation} which others have shown to be effective, we believe the following features to be useful (and humanly justifiable), and offer an investigation into their use alongside different traffic types:
%?? We use these features, and why...

\fakepara{Global state ($\mathbf{4\cdot{}}$Load)}
these are the existing set of load measurements introduced in \cref{sec:environment-and-rl-algorithm}.
These indicate the overall health of the network, and crucially are all measurements which an agent directly controls.

\fakepara{Source IP address}
while ordinarily trivial to spoof, in real-world scenarios the addresses of a set of reflector nodes might exhibit similarity, most likely through geography.

\fakepara{Last action taken}
this encodes an agent's current belief in the maliciousness of a flow.
This feature also potentially allows forgiveness, serving as a reference point for determining whether a source mistakenly marked as malicious exhibits different falloff behaviour after punishment.

\fakepara{Flow duration and size}
features which describe the length of time a connection has been active, and the amount of data transferred within that time.
An extraordinarily long flow, having sent a lot of data, could be more likely to be an amplifier.

\fakepara{Correspondence ratio}
the ratio between upstream and downstream traffic for a source IP.
We define this to be $C_X = \min(\mathit{up}, \mathit{down})/\max(\mathit{up}, \mathit{down})$, where a value close to 0 indicates strong asymmetry.

\fakepara{$\mathbf{\Delta}$ Send/receive rate}
the change in traffic rates caused by the last action.
Behavioural changes induced by bandwidth expansion/reduction are expected to be most visible in these fields.

\fakepara{Mean inter-arrival time (IAT)}
a measure of how often packets arrive at the agent's parent switch; low IATs indicate a high number of packets per second, and can be a possible marker of malicious behaviour.
We only make use of the mean IAT of \emph{inbound} traffic.

\fakepara{(Per-window) packet count}
the amount of packets sent to/from a source over a flow's lifetime (or the current window of measurement), similar in use to flow size and mean IAT.

\fakepara{Mean packet size per-window}
legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other classes of flooding attack.

The exclusion of features such as source/destination ports or protocol numbers is a deliberate choice.
If \emph{QUIC} (or a similar protocol) were to become ubiquitous, then these fields would have little to no correlation with the class of traffic a flow might contain.
Our aim was to design around this constraint as a form of future-proofing.

\begin{table}
	\centering
	\caption{Tile coding windows for each feature.\label{tab:codings}}

@@

All of the above features, save for global state, are 1-dimensional.
Each is tile-coded with 4 tilings and 6 tiles per dimension, using the windows described in \cref{tab:codings}.
We chose to take a separate tile-coding over each feature, ignoring combinatorial effects at present to minimise the amount of extra work an agent must perform.
Indeed, combining too many features into one tile-coding can lead to poor performance due to an inability to generalise action values across similar states.
On the other hand, treating each individually allows us to achieve a quick understanding of the importance of each while seeing adequate performance.

\Cref{fig:udp-feature-plots} shows the effectiveness of each for UDP (resp.\ \cref{fig:udp-feature-plots} for TCP), using a topology with $n=2$ hosts per egress point.
%?? Core findings---different protocols need different features, so everything we proposed above has a use!
The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
%?? How do they do when combined after individual training? Pretty well, especially for TCP.
Additional testing shows that the learned policies may be effectively combined (by summing action values), and are particularly effective for TCP; these results are omitted to preserve space.

%?? NEED TO SAY SOMEWHERE: Baselines use TC for establishing interleaving/buffers filling etc., when MARL+ or SPF are in play then disabling TC (thus allowing illegal states, with \num{-1} reward marking them as such---``negative reinforcement'') may accelerate learning. We have a choice---but the results show that for TCP we WANT TC enabled... For UDP, we actually get a slight degradation in performance. This suggests a kind of ``banding'', perhaps? Important to mention that it easily learns how to maintain a logical bandwidth limit.
During training of the above feature models, we elected to remove the physical bandwidth limit to allow agents to visit illegal states and receive \emph{negative reinforcement} for doing so.
It is important to note that this did not affect the agents' ability to keep the total load below $U_s$, as this behaviour can be easily learned in all cases.
%?? talk about cap difference: \cref{fig:tcp-cap-feature-plots}---we want cap enabled for TCP.
We found this led to a marginal improvement in efficacy when UDP was the primary source of benign traffic, but observed that this hindered learning when TCP was the primary protocol under protection (\cref{fig:tcp-cap-feature-plots})---likely due to the loss of key system dynamics.
It is anticipated that this difference might prove problematic for heterogeneous traffic mixtures one way of solving this may be to investigate the effect of replacing $U_s$ with e.g., $0.99 U_s$ in \cref{eqn:reward-but-better} to introduce a small band of negative reinforcement without actually removing the load restriction in the environment.
Our intent is to potentially accelerate learning while preserving the effects of system dynamics---we term these \emph{banded} rewards.

%\subsection{What's done with these?}

%?? Tile coded together, action taken per-flow every timestep.

%?? NOTE: think about how I might need to change/alter the tile-coding when we add extra discriminative features... (i.e., are there better ways to handle covariance between properties of system/flow state, properly capture generalisation?)

%?? What other consequences are there? We're training each agent on all the flows it's acting on which means slight theoretical changes (we have multiple experience traces, and each uses its last state + new state + new target value (reward) as an update to the \emph{current model}).

@@

?? How to organise? by traffic type, or by model?

?? Probably best just to look at top level stuff, and THEN a simplified comparison for each intended improvement (like banded rewards).

@@

\fakepara{DDoS Prevention}
\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.

@@

An \emph{indirect-control} RL problem, where the role played by the agents is to act in service to \emph{another technique} responsible for decision-making, further optimising or generalising aspects of its operation beyond that of hand-coded heuristics.
A past example includes learning when best to \emph{communicate} and share knowledge between \emph{hidden Markov model} anomaly detectors \cite{DBLP:conf/paisi/XuSH07}.
The position of this work is weakened by its reliance on the discredited `DARPA99' dataset \cite{DARPA-IDD, DBLP:conf/cisda/TavallaeeBLG09, DBLP:conf/sp/SommerP10}, but the idea itself is well-treated and this acts as a driver for improvements in this direction.
Outside of anomaly/intrusion detection, there has been growing interest in the use of reinforcement learning in data-driven networking, such as for intra-AS route optimisation \cite{DBLP:conf/hotnets/ValadarskySST17} and for resource-constrained process allocation \cite{DBLP:conf/hotnets/MaoAMK16}.
Work by \textcite{DBLP:conf/sigcomm/MaoNA17} employs client-side observations of network state and video player performance to optimise adaptive bitrate selection for multimedia streaming by RL methods.

\emph{AuTO} \cite{DBLP:conf/sigcomm/ChenL0L18} employs deep RL to perform traffic optimisation, using the \emph{deterministic policy gradients} algorithm \cite{DBLP:conf/icml/SilverLHDWR14}.
Crucially, they find that the vast majority of flows are short-lived, requiring effective decisions in less than a millisecond.
To overcome the high latency of action computation via neural network, two agents are trained, handling aspects of short and long flows respectively.
The first learns to optimise the flow size thresholds which define priority allocation at each switch (and accordingly learn how best to demarcate long and short flows); these short flows are routed by ECMP.
The second agent makes bespoke decisions about routing, prioritisation etc.\ for each of the remaining long flows, who are likely to live long enough that any actions will have a significant impact.

@@

?? Is the state space interpretable? Yes!
